---
title: "Designing a Hawkin Capture-Style Performance Dashboard"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Designing a Hawkin Capture-Style Performance Dashboard}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
dependencies:
  - knitr
  - shiny
  - bslib
  - plotly
  - dplyr
  - tidyr
  - reactable
  - shiny.i18n
  - lubridate
  - purrr
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Purpose

World-class athletes and sport scientists need a unified place to review, interpret, and act on Hawkin Dynamics force-plate data. This vignette sketches a Shiny dashboard that emulates **Hawkin Capture** while adding:

- **Full data coverage:** force-time curves, jump summaries, readiness flags, and wellness context.
- **Visual clarity:** interactive charts, session timelines, and positional benchmarks.
- **Automations:** scheduled pulls from `hawkinR`, proactive alerts, multilingual UI, and role-aware views.

The goal is to provide a production-ready blueprint you can adapt to your organization's infrastructure.

## Data pipeline

A repeatable pipeline keeps the dashboard fresh and trustworthy:

1. **Authentication:** request an access token with `get_access()` and store it in a secure key store (for example, environment variables or an encrypted file created via [Store_API_Key_Locally](Store_API_Key_Locally.html)).
2. **Ingestion cadence:** run a scheduled job (cron, `taskscheduleR`, GitHub Actions, or Posit Connect) that fetches new data every few minutes while also running a full historical sync daily.
3. **Data model:** persist tests, athletes, and derived metrics in a relational store (Postgres/SQLite) or Parquet files. Use stable IDs (`testId`, `athleteId`, `testTypeId`) for joins and change tracking.
4. **Derived features:** compute jump strategy indicators (RSI, peak power, asymmetry), freshness scores, and individualized z-scores against rolling windows.
5. **Validation:** log ingestion summaries (rows inserted/updated, error rate) and surface them in an admin panel so staff can spot issues early.

### Run it directly

You can execute a minimal end-to-end pull without wiring up your own scheduler. The repo ships with a runner that caches tests and a subset of force-time curves, then optionally launches a Shiny preview:

```r
source(system.file("extdata", "hawkin_capture_runner.R", package = "hawkinR"))
cache <- hawkin_capture_pipeline(days = 7, max_forcetime = 15)
preview_hawkin_capture(cache) # requires shiny + plotly
```

The script reads `HAWKIN_REFRESH_TOKEN`, `HAWKIN_REGION` (default `Americas`), and optional `HAWKIN_ORG_NAME` environment variables before calling `get_access()`, `get_tests()`, and `get_forcetime()`. Output is saved in `hawkin_capture_cache/` to keep downstream rendering fast.

### Example ingestion script (scheduled)

```{r ingest, eval = FALSE}
library(hawkinR)
library(dplyr)
library(lubridate)
library(DBI)
library(duckdb)

con <- dbConnect(duckdb::duckdb(), "hawkin.duckdb")

get_access(
  refreshToken = Sys.getenv("HD_REFRESH"),
  region = Sys.getenv("HD_REGION")
)

from_ts <- floor_date(Sys.time() - minutes(15), unit = "minute")
to_ts   <- Sys.time()

recent_tests <- get_tests(from = from_ts, to = to_ts)
force_traces <- purrr::map_dfr(recent_tests$testId, get_forcetime, .id = "testId")

# Example: compute quick readiness score (simple placeholder)
recent_tests <- recent_tests |>
  mutate(
    peak_power_rel = peakPower / bodyWeight,
    rsi_mod = jumpHeight / timeToTakeoff,
    readiness_score = scales::rescale(peak_power_rel + rsi_mod, to = c(0, 100))
  )

# Persist
if (!dbExistsTable(con, "tests")) dbWriteTable(con, "tests", recent_tests)
else dbAppendTable(con, "tests", recent_tests)

if (!dbExistsTable(con, "force")) dbWriteTable(con, "force", force_traces)
else dbAppendTable(con, "force", force_traces)

invisible(dbDisconnect(con, shutdown = TRUE))
```

## UI/UX blueprint

The following layout uses `bslib` for modern theming and `shiny.i18n` for multilingual copy (you can import translations from Hawkin language packs to support staff and athletes worldwide).

- **Global shell:** navigation bar with athlete search, date range, and language selector.
- **Hero tiles:** readiness score, asymmetry alert, session count, and peak power trend with sparkline.
- **Main panels:**
  - **Readiness overview:** percentile rings, individualized z-scores, and fatigue flags.
  - **Force-time explorer:** interactive overlay of trials (Plotly), with toggles for normalized vs. absolute curves.
  - **Jump summary table:** sortable Reactable with conditional formatting for performance tiers.
  - **Wellness context:** questionnaire scores and acute:chronic workload ratio overlays.
  - **Coach notes:** rich text area that writes back to your performance database.
- **Automation cues:** banners for ingestion health and a download button for CSV/Parquet exports.

### Minimal Shiny scaffold

```{r shiny-skeleton, eval = FALSE}
library(shiny)
library(bslib)
library(plotly)
library(reactable)
library(dplyr)
library(tidyr)
library(shiny.i18n)

# Load translations exported from the Hawkin language pack
translator <- Translator$new(translation_json_path = "inst/i18n/translations.json")
translator$set_translation_language("en")

theme <- bs_theme(bootswatch = "cosmo", primary = "#1E88E5", secondary = "#43A047")

dashboard_ui <- page_navbar(
  title = "Hawkin Capture+",
  theme = theme,
  nav_spacer(),
  input_switch("language", translator$t("Language"), value = TRUE,
               on_label = "EN", off_label = "ES"),
  input_date_range("date", translator$t("Date range")),
  nav_panel(
    translator$t("Overview"),
    layout_columns(
      col_widths = c(3, 3, 3, 3),
      value_box(translator$t("Readiness"), textOutput("readiness")),
      value_box(translator$t("Asymmetry"), textOutput("asym")),
      value_box(translator$t("Sessions"), textOutput("sessions")),
      value_box(translator$t("Peak Power"), plotOutput("pp_spark"))
    ),
    card(
      full_screen = TRUE,
      card_header(translator$t("Force-Time Explorer")),
      plotlyOutput("force_plot"),
      sliderInput("smooth", translator$t("Smoothing (ms)"), 0, 25, 5, step = 5)
    ),
    card(
      card_header(translator$t("Jump Summary")),
      reactableOutput("jump_table")
    )
  ),
  nav_panel(
    translator$t("Wellness & Notes"),
    layout_columns(
      card(plotlyOutput("wellness")),
      card(textAreaInput("notes", translator$t("Coach notes"), width = "100%"))
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$language, translator$set_translation_language(ifelse(input$language, "en", "es")))

  # Replace with data from your ingestion store
  tests <- reactive({ dplyr::tibble(athlete = "A. Smith", jumpHeight = 0.45, asym = 4.2) })

  output$jump_table <- renderReactable({
    reactable(tests(), columns = list(
      jumpHeight = colDef(name = translator$t("Jump Height (m)"), format = colFormat(digits = 2)),
      asym = colDef(name = translator$t("Asymmetry (%)"), format = colFormat(digits = 1),
                    style = function(value) ifelse(value > 10, list(color = "#E53935"), NULL))
    ))
  })
}

shinyApp(dashboard_ui, server)
```

## Analytics and interpretation

- **Force-time:** highlight impulse-to-100 ms, concentric peak power, and net braking/loading symmetry.
- **Jump strategy:** plot time-to-takeoff vs. jump height to classify strategies (dip, stretch-shortening, squat jump drift).
- **Fatigue/readiness:** rolling z-scores (7–21 days) for RSI-mod and peak power; flag green/amber/red bands.
- **Return-to-play:** asymmetry thresholds by limb and task (CMJ, ISO-MID, isometric calf); include smallest worthwhile change bands.
- **Comparative lenses:** team, position group, and individual overlays; automatically pick priors based on athlete profile.

## Automation and alerting

- **Proactive notifications:** email/Slack/Teams when readiness drops \>1 SD from baseline or when asymmetry \>10% for two consecutive sessions.
- **Data quality:** send alerts if ingestion latency \> 10 minutes or if incoming tests fall below expected volume for the day.
- **Export jobs:** nightly Parquet/CSV packages with metadata so coaches can conduct offline analysis.

## Operational checklist

- Version control the Shiny code and the translation JSON used by `shiny.i18n`.
- Containerize the app (Rocker + system libraries) and deploy on Posit Connect, Shiny Server, or Kubernetes behind SSO.
- Log performance: chart render times, API call durations, and cache hit rates to keep the UI under 1.5 seconds for primary views.
- Add accessibility: ARIA labels on controls, high-contrast theme toggle, and keyboard navigation for athletes with visual strain.

By combining `hawkinR` ingestion with this dashboard scaffold, practitioners can interact with Hawkin Dynamics data at the speed required for elite performance, while multilingual support keeps the experience consistent across global staffs.
